\chapter{Introduction}
\label{intro}

In the last years, the field of computer science is permanently shaken by one 
term: Artificial neural networks (\zk{ANN}s). Some people perceive it almost
as a magical formula. And even though \zk{ANN}s are not a spell able to solve
everything, they have wide applications. Their applications are in finance,
data mining, language recognition, computer vision and many more.

Another term that can be heard more and more is \textit{the information age}.
The~availability of computers and the growth of memory limits result in huge
amounts of data.

The huge amount of data is a fuel for \zk{ANN}s. One hundred years ago, an
artificial intelligence was a phantasmagoria of science-fiction writers. Fifty
years ago, it was an idea facing only derision. Twenty-five years ago, a design
doomed to failure due to a lack of training data. Twelve years ago, a bold idea
of few men. Five years ago, an earthquake in each branch of the field of
computer science.

The data availability and their opening in the field of geomatics, as well as
the~computer performance, open doors for the usage of \zk{ANN}s in geographic
information systems (\zk{GIS}). Results of a special type of \zk{ANN}s called
convolutional neural networks (\zk{CNN}s) promises a lot in computer vision and
therefore also in \zk{GIS}, in tasks of detection and classification.

Chapter \ref{cnn} will briefly introduce the~theory behind \zk{CNN}s. In the
first part, the~history and the motivation behind them will be described. The
second part covers multiple types of layers used in \zk{CNN}s. Few pioneering
architectures will be covered.

Chapter \ref{image-ann} will introduce the term computer vision. Various tasks
of computer vision will be named and few breakthrough architectures not 
mentioned in the \zk{CNN} chapter will be described there. The research which 
concluded in the selection of Mask \zk{R-CNN} as the architecture for the 
implementation will be depicted in this chapter.

Chapter \ref{technologies} will describe some of the most important 
technologies used during
the~above-mentioned implementation.

The implementation will be the topic of chapter \ref{implementation}. It will 
summarize the motivation behind the architecture choose and code decisions, and
then describe the~uppermost parts of the code. The practical part of the~thesis
is exactly this implementation.
